A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae

### 2021-03-19 thesis_chapter_3.Rmd
updates to rmd files

---
title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown. Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumlation zones for solitary floating macroalage."
---

@@

In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations. It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain. The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectivley. Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies. This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure. The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory. In addition, the modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface. Therefore, separate simulations will be run to compare the entrained characteristics of passive particle with a kelp particle.

@@

The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity. These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine. The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring. In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

@@

Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'. All particles that have any form of drag exposure are hereafter referred to as 'kelp' particles. All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018]. Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018]. To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios. These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag. These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation. For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively. This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018]. The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1). Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation. The particles sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion. Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations. It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain. This was not considered a concern, as once again, the aim was to investigate surface drift behavior. The time period chosen to run the simulations was 01-01-2018 un 30-01-2018. This particular time period was chosen as the splitting of the Benguela Jet occurs in Janurary, which also conincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].

```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap=" Details for each simulation run for the study. The table cosists of parameterisations for each scenario. The ID coloumn refers to the syntax used in the relevant OceanParcels script (see additional material for link to script). Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure. The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated."}

@@

sim_table <- flextable(data = sim_table)
sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
sim_table <- align(x= sim_table, align = "center", part = "all")
sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
sim_table

@@

@@

Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com). The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean. The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024). The model is forcedhas a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid. In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004). The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$. Both averaged ocean current and wind data are stored as meridonal and zonal velocity vectors.

@@

The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field. Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors, as per @furnans2008. Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step. The same approach was used for the wind drag force. The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.

@@

The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020]. The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 1). The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.

@@

```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap=" Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation. Approximate shape refers to the dimensions of the geometric shape used to represent a particular morpholgical characteristic. Equations used are the equations relating to the dimensions used for the geometric shape estimate. Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas. The total surface areas for each plant characteristic for each surface area type make up the remaining columns."}

@@

area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
area_table <- flextable(data = area_table)
area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
area_table <- align(x= area_table, align = "center", part = "all")
area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
area_table <- set_caption(area_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
area_table
```
## Analysis

In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced. Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018]. In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs. The distance that particles traveled was also measured and used to produce boxplots. Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particle [@VanSebille2018]. If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].

@@

The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A). When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation. The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B). The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site. When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident. In all simulations particles flowed in a north-westward direction.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer). The plots are regular hexagons within the density of particles is calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}

@@

```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder. Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylinderical shape coefficient. The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}

@@

Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag. In addition, not only are the trajectories similar but so are the end points at the end of simulation. However, when *ad-hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories. The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location. The difference in trajectory and distance covered can be attributed to the inclusion of wind drag. Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other. However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020]. This causes the kelp particles to travel further as well as exposed to different time varying flows. Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020]. A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*. The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momemtum energy transfer from wind. This is also reflected in the analysis comparing distances travelled by all the particles in each simulation. The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag.

@@

The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter thier overall trajectory. Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridonal velocity vectors. Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy. The entrainment and expulsion from eddies and vorticities are an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, these studies investigated large rafts and not floating individuals. Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.

The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations. These results cleary show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain. Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy. In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found. The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a vorticy. The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory. Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field. Therefore, more accurately parameterizing shape will therefore allow for further improvements in models of macroalgal dispersal patterns.

>>>>

---
title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown. Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not considered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectories of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumulation zones for solitary floating macroalage."
---

@@

In this study, a numerical approach is used to shed light on the role of hydrodynamic and direct wind drag on macroalgal drift trajectory using Lagrangian based trajectory simulations. It should be noted that this study does not attempt to characterise the dispersal of drifting macroalage within the study domain. The authors argue that too little is know about the drift characteristics of solitary floating macroalgae to characterise dispersal effectively. Instead, this study will aim to shed light on the 'behaviour' of surface drifting macroalgae from a coastal environment into the open ocean, which will inform future characterisation studies. This will be achieved by comparing the trajectory outputs of virtual 'passive' particles (no drag characteristics added) with kelp particles (drag characteristics added) of varying degrees of hydrodynamic and wind drag exposure. The different degrees of wave and wind exposure are meant to reflect the different buoyancy characteristics, as well as identify the role of magnitude of the drag components on overall trajectory. In addition, the modeling of single kelp offers the opportunity to study the role of eddies and vortices in accumulation points on the surface. Therefore, separate simulations will be run to compare the entrained characteristics of passive particle with a kelp particle.

@@

The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity. These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017], generally flowing northward along the coastline and begin to move offshore near Cape Columbine. The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, Eckman transport and geostrophic currents [@stramma1989; @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring. In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2005; @blanke2002] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm Agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

@@

Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'. All particles that have any form of drag exposure are hereafter referred to as 'kelp' particles. All simulations have the same release location (Kommetjie) and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018]. Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018]. To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios. These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag. These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation. For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively. This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018]. The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1). Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation. The particle sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion. Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations. It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain. This was not considered a concern, as once again, the aim was to investigate surface drift behavior. The time period chosen to run the simulations was 01-01-2018 un 30-01-2018. This particular time period was chosen as the splitting of the Benguela Jet occurs in January, which also coincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019]. A summary of the parameters for each simulation are provided in table 1.

```{r sim table, echo=FALSE, message=FALSE, warning=FALSE}

@@

sim_table <- flextable(data = sim_table)
sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
sim_table <- align(x= sim_table, align = "center", part = "all")
sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
sim_table <- set_caption(sim_table, "Table 1: Details for each simulation run for the study. Type refers to either a passive particle with no drag behaviour or a kelp particle that has been advected with any drag exposure. The drag exposure column refers to the parameterisations regarding the level of water and wind drag simulated.")
sim_table

@@

Additional simulations were done for months in 2018, which were chosen based on seasonality. These simulations had the same parameters used for January and were performed for April (Autumn), July (winter), and October (Spring) in order to assess if any patterns seen in January translate into general patterns for times periods throughout 2018.

@@

Both passive and kelp particles were simulated around the South African coastline within hind-cast outputs from the Copernicus Marine Environment Monitoring Service (Copernicus, http://marine.copernicus.com). The Copernicus model is based on the modelling framework known as Nucleus for European Modelling of the ocean. The Copernicus outputs used in this study contains 3D daily current information from the top layer to the bottom (Global Analysis Forecast PHY_001_024). The model is forcedhas a spatial resolution of 0.08$^\circ$ and is interpolated on an Arakawa C native grid. In order to incorporate the effects of direct wind drag within the simulations, the Copernicus Marine Environment Monitoring Service global wind product was used (Global Ocean Wind L4_NRT_012_004). The outputs from the model is composed of 6-hourly averaged fields of surface 10m wind velocity vectors with a spatial resolution of 0.25$^\circ$. Both averaged ocean current and wind data are stored as meridional and zonal velocity vectors.

@@

The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field. Hydrodynamic and wind drag components were calculated separately for both the meridional and zonal velocity vectors, as per @furnans2008. Since drag force is dependent on the velocity vectors which vary with time, the meridional and zonal velocities were interpolated and used in the drag force calculation for each time-step. The same approach was used for the wind drag force. The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.

@@

The cross-sectional area calculated was site specific and the dimensions needed were garnered from morphology data from a previous study [@coppin2020]. The minimum, mean and maximum were calculated for each morphological characteristic, which were used for calculating the overall minimum, mean and maximum cross-sectional areas needed to run the various simulations (see table 2). The trajectory of these 'types' of cross-sectional areas (minimum, mean and maximum overall cross-sectional area) were used to determine the influence of drag, both hydrodynamic and wind, on overall trajectory.

@@

```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE}

@@

area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
area_table <- flextable(data = area_table)
area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
area_table <- align(x= area_table, align = "center", part = "all")
area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
area_table <- set_caption(area_table, "Table 2: Table summarising the attributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation. Approximate shape refers to the dimensions of the geometric shape used to represent a particular morphological characteristic. Equations used are the equations relating to the dimensions used for the geometric shape estimate. Plant dimensions are the measurements used, collected in previous work by @coppin2020, used to calculate the associated surface areas. The total surface areas for each plant characteristic for each surface area type make up the remaining columns.")
area_table
```
## Analysis

In order to compare trajectories between cross-sectional area types and wind exposure scenarios, mean trajectories were calculated and density distribution maps were produced. Density distribution maps are an effective method for assessing pathways and ocean connectivity [@vanSebille2018]. In addition, mean trajectory was calculated and included in the density distribution maps to further aid in comparing simulation outputs. The distance that particles traveled was also measured and used to produce boxplots. Distance traveled from the release site can help reveal topographic steering and momentum energy transfer from the wind field to kelp particles [@vanSebille2018]. If the wind direction is in the same direction as the current the virtual particles will cover larger distances due to an increase in the velocity as a result of momentum energy transfer [@putman2020; @putman2018; @vanSebille2018].

@@

The mean trajectories of the different cross-sectional areas are similar to the mean trajectory of the passive particles (Figure 2A). When comparing mean trajectories between the different cross-sectional areas for the 100% hydrodynamic drag simulations, no differences can be seen; rather, there is only a slight variation. The results from the simulations including different hydrodynamic and wind drag exposures and cross-sectional area types show differences in mean trajectory when compared to the passive simulation (Figure 2B). The end points of the simulations including hydrodynamic and wind drag were different compared to the passive simulations, as well the final location of the kelp particles being further from the release site. When comparing the different hydrodynamic and wind drag simulations to each other, only slight variation in mean trajectories was evident. In all simulations particles flowed in a north-westward direction. The patterns of north-westward flow and entrainment in eddies is also evident in the additional simulations, however clear differences in mean trajectory between kelp and passive particles is clear (see appendix).

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.cap='Figure 3: Comparison of density plots at the end run time for scenarios 1 - 13, for January 2018 (summer). The plots are regular hexagons within the density of particles was calculated. The mean trajectories for each scenario are also plotted for comparison. The reader is referred to tables 1 and 2 for details pertaining to each scenario.', fig.height=10.2, fig.width=7.5}

@@

```{r shape data outputs, warning=FALSE, echo=FALSE, message=FALSE, fig.cap="Figure 6: Comparison of trajectory data between a kelp sphere and kelp cylinder. Plot A is the comparison of mean trajectories, plot B is the comparison of total distance travelled by all particles, and plot C and D are the density plots for simulations using a sphere and cylindrical shape coefficient. The reader is referred to tables 1 and 2 for details pertaining to each scenario.", fig.height=8, fig.width=8}

@@

Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow the prevailing surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag. In addition, not only are the trajectories similar but so are the end points at the end of simulation. However, when *ad-hoc* wind drag (windage) is considered in combination with hydrodynamic drag the end locations differ greatly when compared to purely Lagrangian particle trajectories. The inclusion of wind drag causes the kelp particles to flow further away from the coast and with a greater distance travelled from the release location. The difference in trajectory and distance covered can be attributed to the inclusion of wind drag. Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other. However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object [@hackett2006; @putman2020]. This causes the kelp particles to travel further as well as exposed to different time varying flows. Past studies have shown this when including windage as a function of inertia [@putman2018;@brooks2019; @putman2020]. A study by @putman2020 used *ad-hoc* windage factors that was based on the surface current velocity to assess the approach for improving transport predictions of pelagic *Sargassum sp*. The results showed that including ad-hoc windage factors improved virtual trajectories compared to tracked *Sargassum sp* mats which was partly due to the inclusion of momentum energy transfer from wind. This is also reflected in the analysis comparing distances travelled by all the particles in each simulation. The particles in the simulations that included wind drag traveled significantly greater distances compared to the passive or hydrodynamic only drag. Similar patterns are seen in the additional simulations run for April (autumn), July (winter), and October (spring), however even kelp particles with only hydrodynamic drag flow further compared to passive particles in these times of the year. This may be a result of the kelp particles exposed to different time-varying flows, which results in the kelp particles following a different surface current vector compared to passive particles. This also results in different end-points and therefore these results further confirm that drag is an important factor to account for when investigating macroalgal drift patterns.

@@

The ocean is comprised of time-varying flows and changes in an objects velocity that can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity in trajectory between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchling trajectories are mostly influenced by the ocean currents in the first year of hatching due to  as juveniles do are not swim fast enough to alter their overall trajectory. Therefore, the differences in cross-sectional area types may not be significant enough to have an effect on the relevant zonal and meridional velocity vectors. Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of Lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of Brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggest that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a  high density of particles entrained in a eddy. The entrainment and expulsion from eddies and vortices are an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, these studies investigated large rafts and not floating individuals. Inertia is based on the size/mass of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles, and will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.

The role of eddies and vortices in the entrainment of kelp is further supported by the accumulation simulations. These results clearly show that the inclusion of drag greatly increases the number of particles entrained in the eddy within the study domain. Eddies and vortices may act as accumulation zones for kelp where the formation of rafts may occur, or the kelp is entrained for long enough to eventually sink in area below the eddy. In a study by @woodborne1989 references a site offshore approximately 120km from Cape Columbine where dredged kelp was found. The area referenced by @woodborne1989 coincides with the area where virtual kelp are entrained by a eddy. The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory. Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field. Therefore, more accurately parameterizing shape will therefore allow for further improvements in models of macroalgal dispersal patterns.
